---
title: "Homework 2 Practice"
author: "ASHLEY ROMO"
date: "2023-09-29"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1
```{r clean_538_pols}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("hw2_data/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```



```{r clean_538_snp}
snp = 
  read_csv("hw2_data/snp.csv") |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

Finally, we tidy the `unemployment` data so that it can be merged with the `pols` and `snp` datasets.

```{r clean_538_unemp}
unemployment = 
  read_csv("hw2_data/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Now we merge the three datasets!

```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```

# Problem 2
```{r}
mr_wheel_df =
  read_xlsx("hw2_data/202207 Trash Wheel Collection Data.xlsx",1, range = "A2:N549") |> 
  janitor::clean_names()|> 
  mutate(homes_powered = weight_tons*500/30
    )

prof_wheel_df =
  read_xlsx("hw2_data/202207 Trash Wheel Collection Data.xlsx",2 , range = "A2:M96") |> 
  janitor::clean_names()|> 
  mutate(homes_powered = weight_tons*500/30
    )

gwyn_wheel_df =
  read_xlsx("hw2_data/202207 Trash Wheel Collection Data.xlsx",4 , range = "A2:K108") |> 
  janitor::clean_names()|> 
  mutate(homes_powered = weight_tons*500/30
    )


merged_df =
  full_join(mr_wheel_df, 
            prof_wheel_df,
            gwyn_wheel_df, 
            by = "dumpster")

```

## Problem 3

```{r}
mci_df = 
  read_csv("hw2_data/MCI_baseline.csv", skip = 1) |> 
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex, 
      1 ~ "male",
      0 ~"female"
    ),
    apoe4 = case_match(
      apoe4, 
      1 ~ "carrier",
      0 ~ "non-carrier"
    ), 
  ) |> 
   select(id, baseline_age = current_age, onset_age = age_at_onset, everything())
```

An important step in the import process include skipping the first row in the csv file because it is a legend about the dataset and not data. Relevant features of the dataset include the age at baseline, age at onset, and the whether the participant is a carrier or non-carrier of the apoe4 variant. These are relevant because they provide information to investigate whether participants with the apoe4 variant develop MCI at a younger age.

```{r}
# number of participants at baseline 
nrow(mci_df)

# participants who developed MCI
developed_df = 
  filter(mci_df, onset_age != ".")


# average age at baseline
mean(pull(mci_df, baseline_age))

# proportion of women with apoe4 variant 
female_car_df = 
  select(mci_df, sex, apoe4) |> 
  filter(sex == "female", apoe4 == "carrier")

```

There were a total of `r nrow(mci_df) ` participants recruited at baseline. 

Of the total participants at baseline, `r nrow(developed_df)/nrow(mci_df)` was the proportion of participants who developed MCI.

The average age at baseline is `r mean(pull(mci_df, baseline_age))`. 

The proportion of women in the study who are carriers is `r nrow(female_car_df)/nrow(mci_df)`.



